---
title: "Gov 1006 Final Project - Milestone 4"
author: "Suriya Kandaswamy"
output: 
  bookdown::pdf_document2:
    keep_tex: true
    biblio-style: "apalike"
header-includes:
   - \usepackage{dcolumn} 
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(gt)
library(stargazer)
library(tidyverse)
library("rstanarm")
```

# Introduction
The paper _Policy Preferences and Policy Change: Dynamic Responsiveness in the American States, 1936-2014_ [@caughey2018policy] discusses the predictors of state policy change and pinpoints what factors have the strongest impact on a state's liberalism. In particular, the authors focus on mass political consensus and state policy response to changes in this consensus. They have acquired measures of mass policy preferences through an accumulation of multiple poll responses over time and of state policy liberalism using some defined estimates ^[These quantified liberalism estimates are estimated following a procedure defined by @caughey2016dynamics]. Responsiveness is then measured as a relationship between the liberalism of the masses and the liberalism of the policies at each given moment in time.

# Milestone Exercise
For the purposes of this milestone, we'll be using the dataset __countrypops__ that is built into the __gt__ package. All code can be found in my Github repository ^[https://github.com/sardination/gov1006-final-project].

The following table shows population growth over time in three North American countries - Canada, United States, and Mexico.
```{r display_table}
# Generate table of populations of three North American countries
min_year <- min(countrypops$year)
max_year <- max(countrypops$year)

northamerica <- countrypops %>%
  dplyr::filter(country_code_3 %in% c("CAN", "USA", "MEX")) %>%
  dplyr::filter(year %% 10 == 0 | year %in% c(min_year, max_year)) %>%
  dplyr::mutate(Year = year) %>%
  dplyr::select(country_name, Year, population) %>%
  pivot_wider(names_from = country_name, values_from = population)

northamerica %>%
  gt() %>%
  tab_header(
    title = "North American Population Over Time",
    subtitle = paste(min_year, "-", max_year)
  )
```

This next table shows the lienar regression fits for the populations of Canada and Mexico based on year and the population of the United States.
```{r regression_table, results='asis'}
can_fit <- lm(Canada ~ Year + `United States`, data = northamerica)
mex_fit <- lm(Mexico ~ Year + `United States`, data = northamerica)

stargazer(can_fit, mex_fit, title="Regression Results", align = TRUE, header=FALSE, type='latex')
```


# References







