---
title: "Gov 1006 Final Project - Milestone 2"
output: html_document
---

## Suriya Kandaswamy

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Policy Preferences and Policy Change: Dynamic Responsiveness in the American States, 1936-2014
#### Devin Caughey and Christopher Warshaw

This paper discusses the predictors of state policy change and pinpoints what factors have the strongest impact on a state's liberalism. In particular, the authors focus on mass political consensus and state policy response to changes in this consensus. They have acquired measures of mass policy preferences through an accumulation of multiple poll responses over time and of state policy liberalism using some defined estimates. Responsiveness is then measured as a relationship between the liberalism of the masses and the liberalism of the policies at each given moment in time.

[Paper](https://www-cambridge-org.ezp-prod1.hul.harvard.edu/core/services/aop-cambridge-core/content/view/8EC04664B89C28D46EDC266E3313BBBC/S0003055417000533a.pdf/policy_preferences_and_policy_change_dynamic_responsiveness_in_the_american_states_19362014.pdf) | [Replication Data](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/K3QWZW)

### Replication Data
Summary:
```{r data_summary}
load("caughey_warshaw_summary.RData") # saves data into `table` variable
summary(table)
```

Glimpse:
```{r data_glimpse}
glimpse(table)
```


Repository: [https://github.com/sardination/gov1006-final-project](https://github.com/sardination/gov1006-final-project)