---
title: "Gov 1006 Final Project - Milestone 6"
author: "Suriya Kandaswamy"
output: 
  bookdown::pdf_document2:
    keep_tex: true
    biblio-style: "apalike"
header-includes:
   - \usepackage{dcolumn} 
   - \usepackage{subfig}
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(gt)
library(stargazer)
library(tidyverse)
library("rstanarm")
```

# Introduction
The paper _Policy Preferences and Policy Change: Dynamic Responsiveness in the American States, 1936-2014_ [@caughey2018policy] discusses the predictors of state policy change and pinpoints what factors have the strongest impact on a state's liberalism. In particular, the authors focus on mass political consensus and state policy response to changes in this consensus. They have acquired measures of mass policy preferences through an accumulation of multiple poll responses over time and of state policy liberalism using some defined estimates ^[These quantified liberalism estimates are estimated following a procedure defined by @caughey2016dynamics]. Responsiveness is then measured as a relationship between the liberalism of the masses and the liberalism of the policies at each given moment in time. All code can be found in my Github repository ^[All analysis for this paper is available in the Github repository at https://github.com/sardination/gov1006-final-project].

Ultimately, the authors of the paper run different types of regressions using mass liberalism (from opinion data) and policy liberalism (using measures as described in @caughey2016dynamics) as independent variables in addition to other explanatory factors such as whether a state is in the "South" or not, whether it was an election year, and a number of other features in order to determine how mass liberalism impacts policy liberalism. They find that state policies tend to respond to mass opinion in terms of liberalism, both for economic and social issues. They also find that state policies respond more strongly in the social case than in the economic one.

Interestingly, these correlations may have become more pronounced since the Carter presidency due to an increase in partisan politics shifting towards hyperpartisanism, as individual identify more strongly with a party's views on both social and economic issues and thus will reflect policy changes that align with these views [@erikson2006public]. Some scholars, like @achen2017democracy, have claimed that mass liberalism has little to no effect on policy liberalism, but the results of this paper indicate the true effect is possibly otherwise.

# Extension

It might be interesting to see whether Democratic control within a state has an effect on the responsiveness of its policy liberalism to mass liberalism. While the paper has examined the effect that Democratic control in a specified year has on policy liberalism itself, it does not examine whether states with more Democratic control, for example, react more to a state's mass liberalism than states without. There are a few different ways to go about examining this effect. Democratic control is normalized onto a 0-1 scale, so we may seek to split data points into two subsets of Democratic control: 0 - 0.5 and 0.5 - 1. We then perform the same analysis of policy responsiveness to mass liberalism and compare the responsiveness of the two subsets, both for economic and social policy. Another way to do this is to include the interaction between Democratic control and mass liberalism in the formula when trying to explain policy liberalism. This way, any corrections given my Democratic control will contribute to the explanation of policy liberalism, adn we can correct for Democratic control as an effect on *responsiveness* rather than just *liberalism*. Ideally, whether a government is under Democratic control or not should not have an effect on its responsiveness in terms of policy liberalism to mass liberalism and we would hope to see a negligible interaction between mass liberalism and Democratic control when trying to explain policy liberalism. It is possible, though, that one party may be more prone to exercising control in state governments in a way that does not reflect the preferences of the people, and this analysis will hopefully be able to expose whether Democratic control dictates how well a government is able to serve the needs of its people.

```{r load_data}
# Load replication resuts
load("code/output-data/Analyze.RData")

StarsToBold <- function (text) {
  out <- gsub("([[:digit:]]+)\\^\\{\\*+\\}", "\\\\textbf{\\1}", text)
  cat(out, sep="\n")
  invisible(out)
}
```

# Appendix

Using the replication data and code provided by Caughey and Warshaw [@caughey_data], all major findings and figures can be replicated and analyzed as follows. Code is modified as needed for the purposes of this assignment and extension. Figures and tables are replicated under sections with the same headers as shown in the paper. While some numbers differ slightly, all results discussed in the paper were replicated.

## Figure 1 - Mass Liberalism by State, 1940-2010
The top half of the figure shows the population's liberalism by state in terms of opinions about economic topics (mass economic liberalism) while the bottom half shows the population's liberalsim by state in terms of opinions about social topics (mass social liberalism).

```{r liberalism_maps}
# Generate US maps of opinion liberalism as done in 6Figures.R. Edited for the purposes
#   of this assignment

summ.map <- merge(summ, map_data("state"), by.x="name", by.y="region")
summ.map <- arrange(summ.map, group, order)

map.yrs <- c(1940, 1975, 2010)

# Create economic liberalism US map colored by opinion data.
## Economic
summ.map %>%
  filter(Year %in% map.yrs) %>%
  ggplot(aes(x=long, y=lat, group=group, fill=MassEcon00_mean)) +
  facet_wrap(~Year, ncol=3) +
  geom_polygon(colour="black", size=rel(.1)) +
  coord_map("polyconic") +
  viridis::scale_fill_viridis(direction=1) +
  theme_clean() +
  guides(fill=FALSE)

## Social
summ.map %>%
  filter(Year %in% map.yrs) %>%
  ggplot(aes(x=long, y=lat, group=group, fill=MassSocial00_mean)) +
  facet_wrap(~Year, ncol=3) +
  geom_polygon(colour="black", size=rel(.1)) +
  coord_map("polyconic") +
  viridis::scale_fill_viridis(direction=1) +
  theme_clean() +
  guides(fill=FALSE)
```

## Figure 2 - Cross-Sectional Relationship between Mass and Government Policy Liberalism, by Region, Era, and Issue Domain
This figure displays the relationship between mass liberal opinion and the liberalism of government policy, both for economic and social issues. It utilizes two models: one for southern statesa and one for non-southern states. Each category of states is displayed on the plots with its own line.

```{r cross_sectional_mass_policy}
# Generate cross-sectional relationship plots as done in 6Figures.R.

## Economic
summ %>%
  group_by(Pre72, StPO, South11) %>%
  summarise(Years = paste0(min(NC(Year)), "-", max(NC(Year))),
            PolicyEcon00_mean = mean(PolicyEcon00_mean, na.rm=TRUE),
            MassEcon00_mean = mean(MassEcon00_mean, na.rm=TRUE)) %>%
  ggplot(aes(x=MassEcon00_mean, PolicyEcon00_mean, label=StPO,
             color=South11, lty=South11)) +
  facet_wrap(~Years, scales="free") +
  geom_smooth(method=lm) +
  geom_text(show.legend=FALSE, alpha=2/3) +
  scale_colour_manual(values = c("black", "grey40")) +
  theme_bw() +
  theme(legend.position="none", plot.title = element_text(hjust = 0.5)) +
  labs(x="Mass Liberalism", y="Government Liberalism", color=NULL, lty=NULL) +
  ggtitle("Economic Policies")

## Social
summ %>%
  group_by(Pre72, StPO, South11) %>%
  summarise(Years = paste0(min(NC(Year)), "-", max(NC(Year))),
            PolicySocial00_mean = mean(PolicySocial00_mean, na.rm=TRUE),
            MassSocial00_mean = mean(MassSocial00_mean, na.rm=TRUE)) %>%
  ggplot(aes(x=MassSocial00_mean, PolicySocial00_mean, label=StPO,
             color=South11, lty=South11)) +
  facet_wrap(~Years, scales="free") +
  geom_smooth(method=lm) +
  geom_text(show.legend=FALSE, alpha=2/3) +
  scale_colour_manual(values = c("black", "grey40")) +
  theme_bw() +
  theme(legend.position="none", plot.title = element_text(hjust = 0.5)) +
  labs(x="Mass Liberalism", y="Government Liberalism", color=NULL, lty=NULL) +
  ggtitle("Social Policies") 
```

## Table 1 - Cross-sectional and dynamic responsiveness, by issue domain and region.
This table details the linear regression analysis done on explaining policy liberalism in year $t$ given the variables of mass liberalism in the year $t-1$, whether a state is in the "South" or not, and policy liberalism in the year $t-1$.

```{r responsiveness_table, results='asis'}
# Copied from replication code in 8Tables.R. Generates table explaining policy liberalism
#   using prior year mass and policy liberalism

stargazer_table <- capture.output(tab1.mods %>%
  stargazer(
    type="latex"
  , header=FALSE
  , float=FALSE
  , se=tab1.moc.se
  , coef=tab1.moc.est
  , t.auto=TRUE
  , p.auto=TRUE ## from standard normal (very close to t when DF = 3,500)
  , omit=c("Year|Constant|^South11South$", "StPO")
  , omit.labels=c("Year $\\times$ South FEs", "State FEs")
  , omit.stat=c("F", "rsq")
  , omit.table.layout = "n"
  , align=TRUE
  , initial.zero=FALSE
  , dep.var.caption="DV: Domain-Specific Policy Liberalism ($t$)"
  , dep.var.labels=c(
      "------------ Social ------------ $|$ ----------- Economic -----------")
  , dep.var.labels.include=TRUE
  , column.labels=c("XS", "FE", "LDV", "DP", "XS", "FE", "LDV", "DP")
  , order=c("MassM1\\_mean$", "MassM1\\_mean:South11South",
            "PolicyM1")
  , no.space=TRUE
  , digits=3
  , column.sep.width = "0.5pt"
  , font.size = "footnotesize"
  , covariate.labels=c(
      "Mass Liberalism$_{t-1}$",
      "Mass Lib$_{t-1}$ $\\times$ South",
      "Policy Liberalism$_{t-1}$")
  ))

stargazer_table %>%
  StarsToBold()
```

## Table 2 - Effect of mass policy preferences and partisanship on partisan turnover
This table shows the effect that mass liberalism and partisan control in year $t-1$ have on partisan control in year $t$.

```{r pref_partisan_on_turnover_table, results='asis'}
# Copied from replication code in 8Tables.R. Generates table to explain partison turnover
#   using mass policy prefs and partisanship from prior years

stargazer_table <- capture.output(tab2.mods %>%
  stargazer(
    type="latex"
  , header=FALSE
  , float=FALSE
  , se=tab2.moc.se
  , coef=tab2.moc.est
  , t.auto=TRUE
  , p.auto=TRUE ## from standard normal (very close to t when DF = 3,500)
  , omit=c("Year", "StPO|Constant")
  , omit.labels=c("Year FEs", "State FEs")
  , omit.stat=c("F", "rsq")
  , omit.table.layout = "n"
  , initial.zero=FALSE
  , align=TRUE
  , dep.var.caption="DV: Democratic Control Index  ($t$)"
  , dep.var.labels.include=FALSE
  , no.space=TRUE
  , digits=3
  , column.sep.width = "0.5pt"
  , font.size = "footnotesize"
  , order=c("MassSocial", "MassEcon", "PID", "Control")
  , covariate.labels=c(
      "Mass Social Lib$_{t-1}$",
      "Mass Econ Lib$_{t-1}$",
      "Mass Dem PID$_{t-2}$",
      "Dem Control$_{t-1}$"
      )
  ))

stargazer_table %>%
  StarsToBold()

```

## Table 3 - Partisan selection and adaptation as mechanisms of dynamic responsiveness.
This table shows the effect that current (year $t$) partisan control as well as year $t-1$'s mass liberalism, whether it was an election year, and policy liberalism have on social and economic policy liberalsm in year $t$. 

```{r selection_adap_responsiveness_table, results='asis'}
# Copied from replication code in 8Tables.R. Generates table explaining policy liberalism
#   as a prior year liberalism and current year partisanship

stargazer_table <- capture.output(tab3.mods %>%
  stargazer(
    type="latex"
  , header=FALSE
  , float=FALSE
  , se=tab3.moc.se
  , coef=tab3.moc.est
  , t.auto=TRUE
  , p.auto=TRUE ## from standard normal (very close to t when DF = 3,500)
  , omit=c("^Year", "StPO|Constant")
  , omit.labels=c("Year FEs", "State FEs")
  , omit.stat=c("F", "rsq")
  , omit.table.layout = "n"
  , align=TRUE
  , dep.var.caption="DV: Domain-Specific Policy Liberalism ($t$)"
  , dep.var.labels=c(
      "-------------- Social --------------- $|$ ----------- Economic -----------")
  , dep.var.labels.include=TRUE
  , order=c("DemControl", "MassM1\\_mean$", "YearAfterHouseElectionFALSE",
            "YearAfterHouseElectionTRUE", "Policy")
  , single.row=FALSE
  , no.space=TRUE
  , initial.zero=FALSE
  , digits=3
  , column.sep.width = "-10pt"
  , font.size = "footnotesize"
  , covariate.labels=c(
      "Dem Control$_{t}$",
      "Mass Lib$_{t-1}$",
      "Mass Lib$_{t-1}$ (No Elec$_{t-1}$)",
      "Mass Lib$_{t-1}$ (Elec$_{t-1}$)",
      "Policy Lib$_{t-1}$")    
  ))

stargazer_table %>%
  StarsToBold()

```

## Table 4 - Moderators of dynamic responsiveness
This table explains social and economic policy liberalism in year $t$ through several factors (all listed in the table) in year $t-1$.

```{r dynamic_resp_moderators, results='asis'}
# Copied from replication code in 8Tables.R. Generates table explaining policy liberalism based
#   on several factors.

stargazer_table <- capture.output(tab4.mods %>%
  stargazer(
    type="latex"
  , header=FALSE
  , float=FALSE
  , se=tab4.moc.se
  , coef=tab4.moc.est
  , t.auto=TRUE
  , p.auto=TRUE ## from standard normal (very close to t when DF = 4,500)
  , omit=c("^Year|^Pre72Pre-1972$|^Pre72Pre-1972:South11South$",
           "StPO|Constant|^South11South$")
  , omit.labels=c("Year FEs", "State FEs")
  , omit.stat=c("F", "rsq")
  , omit.table.layout = "n"
  , align=TRUE
  , dep.var.caption="DV: Domain-Specific Policy Liberalism ($t$)"
  , dep.var.labels=c(
      "-------------- Social -------------- $|$ ----------- Economic -----------")
  , dep.var.labels.include=TRUE
  , single.row=FALSE
  , no.space=TRUE
  , initial.zero=FALSE
  , digits=3
  , column.sep.width = "-10pt"
  , font.size = "footnotesize"
  , order=c("^MassM1\\_mean$",
            "^MassM1\\_mean:Pre72Pre-1972$",
            "^MassM1\\_mean:South11South$",
            "Pre72Pre-1972:South11South$",
            "Suffrage", "Campaign", "Citizen", "LegDays", "PolicyM1")
  , covariate.labels=c(
      "Mass Liberalism$_{t-1}$",
      "Mass Lib$_{t-1} \\times$ Pre-1972",
      "Mass Lib$_{t-1} \\times$ South",
      "Mass Lib$_{t-1} \\times$ Pre-1972 $\\times$ South",
      "Suffrage Restriction",
      "Suff Restrict $\\times$ Mass Lib$_{t-1}$",
      "Contribution Limits",
      "Contrib Limit $\\times$ Mass Lib$_{t-1}$",
      "Citizen Government",
      "Citizen Gov't $\\times$ Mass Lib$_{t-1}$",
      "Legislative Days (Logged)",
      "Leg Days $\\times$ Mass Lib$_{t-1}$",
      "Policy Liberalism$_{t-1}$")
  ))
stargazer_table %>%
  StarsToBold()
```

# References



